New lower bounds for Schur and weak Schur numbers

### 2021-04-16 article/introduction.tex
reorganisation

\par
Finally, we analyze the significant difference between new lower bounds obtain with templates and the former 
lower bounds obtained by computer search and we provide evidence which indicate that the main assumption 
made in those articles removes the optimal partitions from the search space.

@@

\begin{definition}
A partitition is sometimes referred to as a colouring. The colouring associated to a partition \(A_1, ..., A_n\) of
\([\![1, p]\!]\) is the function \(f\) such that \(\forall x \in [\![1, p]\!], x \in A_{f(x)}\). Likewise, the partition 
associated to a colouring \(f\) of \([\![1, p]\!]\) with \(n\) colors is \(\forall c \in [\![1, n]\!], A_c = f^{-1}(c)\).
\end{definition}

>>>>

\par
Finally, we analyze the significant difference between new lower bounds obtained with templates and the former 
lower bounds obtained by computer search and we provide evidence which indicate that the main assumption 
made in those articles removes the optimal partitions from the search space.

@@

\begin{definition}
A partitition is sometimes referred to as a coloring. The coloring associated to a partition \(A_1, ..., A_n\) of
\([\![1, p]\!]\) is the function \(f\) such that \(\forall x \in [\![1, p]\!], x \in A_{f(x)}\). Likewise, the partition 
associated to a coloring \(f\) of \([\![1, p]\!]\) with \(n\) colors is \(\forall c \in [\![1, n]\!], A_c = f^{-1}(c)\).
\end{definition}

### 2021-04-16 article/schur.tex
reorganisation

\qquad In this section, we use Rowley's template-based constructions \cite{RowleyRamsey} in the context of Schur 
numbers. To improve lower bounds for Ramsay's numbers, Rowley introduces special sum-free partitions verifying 
some additional properties which can be extended using a method generalizing Abbott and Hanson's construction 
\cite{AbbottHanson}. Rowley named these partitions "templates", and we keep this name in the entire article. We 
then find new templates and use them to provide new lower bounds for Schur numbers.

@@

Using a SAT solver, we exhibited SF-templates, hence providing lower bound on \(S^+\) and inequalities 
of the type \(S(n+k) \geqslant a S(n) + b\). We have sought templates providing the greatest value of 
\((a, b)\) (for the lexicographic order). When the number of colors excceeded 5, in order to reduce the search space we 
looked for symmetric SF-templates, we assumed that the special color was the last color to appear and we constrained 
the \(m_c\)'s out of being too small. Further details about the encoding as a SAT problem can be found in the
\hyperref[SAT]{SAT section}. \\

Here are the best inequalities on Schur numbers so far (the templates corresponding to the third, fourth and fifth 
inequalities can be found in the appendix):
\[ S(n+1) \geqslant 3S(n) + 1 \]
\[ S(n+2) \geqslant 9S(n) + 4 \]
\[ S(n+3) \geqslant 33S(n) + 6 \]
\[ S(n+4) \geqslant 111S(n) + 43 \]
\[ S(n+5) \geqslant 380S(n) + 148 \]
\[ S(n+6) \geqslant 1140S(n) + 528 \]

The first inequality comes from the original Schur's paper \cite{Schur1917}. The second one is due to
Abott \cite{AbbottHanson}
and the third one to Rowley \cite{RowleyRamsey}. The other ones are new. \\

The first 3 inequalities are optimal. The fourth one is optimal among symmetric SF-templates whose special color is 
the last in the order of apparition (and with a multiplicative factor less than or equal to 118). The fifth one is most 
likely not optimal but should not be too far from the optimal SF-template. 
Finally, the sixth one is obtained by combining (see below) the SF-template of length 380 and the one of length 3. 
Although we could not find a better SF-template with 7 colors, the last inequality is definetely very far from the 
optimal value. One may try to seek better templates by constraining less the search space and by using 
Monte-Carlo methods, as in \cite{Bouzy2015AnAP}. This could be the suject of a future work.\\

We also have a similar theorem where only \(S^+\) is involved.

@@

The previous inequalities give new lower bounds for \(S(n)\) for
\( n \geqslant 9 \). We compute the lower
bounds for \( n \in [\![8,15]\!] \) using the four different inequalities, please notice that the best values for \( n
=8\) and \(n = 13\) were obtained thanks to the first one, found by Rowley. The best lower bounds are highlighted.

\begin{center}
\begin{tabular}{|*{5}{c|}}
    \hline
	\(n\) & 8 & 9 & 10 & 11 \\
	\hline
	\(33S(n-3) + 6 \) & \cellcolor{yellow} 5286 & 17694 & 55446 & 174444\\
	\hline
	\(111S(n-4) + 43 \) & 4927 & \cellcolor{yellow} 17803 & 59539 & 186523\\
	\hline
	\(380S(n-5) + 148 \) & 5088 & 16868 & \cellcolor{yellow} 60948 & \cellcolor{yellow} 203828 \\
	\hline
	\(1140S(n-6) + 528 \) & 5088 & 15348 & 50688 & 182928\\
	\hline
	\hline
	\(n\) & 12 & 13 & 14 & 15 \\
	\hline
	\(33S(n-3) + 6 \) & 587505 & \cellcolor{yellow} 2011290 & 6726330 & 21072090\\
	\hline
	\(111S(n-4) + 43 \) & 586789 & 1976176 & 6765271 & 22624951 \\
	\hline
\(380S(n-5) + 148 \) & \cellcolor{yellow} 638548 & 2008828 & \cellcolor{yellow} 6765288 & \cellcolor{yellow} 23160388 \\\hline
	\(1140S(n-6) + 528 \) & 611568 & 1915728 & 6026568 & 20295948 \\
	\hline
\end{tabular}
\end{center}
Except for 8, 9 and 13, the best lower bounds are obtained thanks to
the third inequality \( S(n+5) \geqslant 380S(n) + 148\). The table
doesn't go any further, but the same inequality allows to improve the
lower bounds for every \( n \geqslant 15 \).

\begin{corollary}
The growth rate for Schur numbers (and Ramsey numbers \(R_n(3)\))  satisfies \(\gamma \geqslant \sqrt[5]{380} \approx 3.28 \).
\end{corollary}
\textsc{Proof :} It is a mere consequence of the inequality \( S(n+5) \geqslant 380S(n) + 148\). As for Ramsey's
numbers growth rate, a lower bound can be found using Schur's one, thanks to \(S(n) \leqslant R_n(3)-2 \)
(see \cite{Schur1917}).

>>>>

\qquad In this section, we use Rowley's template-based constructions \cite{RowleyRamsey} in the context of Schur 
numbers. In order to improve lower bounds for Schur and Ramsey numbers, Rowley introduces special sum-free 
partitions verifying some additional properties which can be extended using a method generalizing Abbott and 
Hanson's construction \cite{AbbottHanson}. Rowley named these partitions "templates", and we keep this name in 
the entire article. We then find new templates and use them to provide new lower bounds for Schur numbers.

@@

We also have a similar construction theorem where only \(S^+\) is involved.

@@

Using a SAT solver, we exhibited SF-templates, hence providing lower bound on \(S^+\) and inequalities 
of the type \(S(n+k) \geqslant a S(n) + b\). We have sought templates providing the greatest value of 
\((a, b)\) (for the lexicographic order). When the number of colors excceeded 5, in order to reduce the search space we 
looked for symmetric SF-templates, we assumed that the special color was the last color to appear and we constrained 
the \(m_c\)'s out of being too small. Further details about the encoding as a SAT problem can be found in \cite{Heule2017}.

\par
Here are the best inequalities on Schur numbers so far (the templates corresponding to the third, fourth and fifth 
inequalities can be found in the appendix):
\[ S(n+1) \geqslant 3S(n) + 1 \]
\[ S(n+2) \geqslant 9S(n) + 4 \]
\[ S(n+3) \geqslant 33S(n) + 6 \]
\[ S(n+4) \geqslant 111S(n) + 43 \]
\[ S(n+5) \geqslant 380S(n) + 148 \]
\[ S(n+6) \geqslant 1140S(n) + 528 \]

The first inequality comes from  Schur's original article\cite{Schur1917}. The second one is due to
Abott \cite{AbbottHanson}
and the third one to Rowley \cite{RowleyRamsey}. The other ones are new.

\par
The first 3 inequalities are optimal. The fourth one is optimal among symmetric SF-templates whose special color is 
the last in the order of apparition (and with a multiplicative factor less than or equal to 118). The fifth one is most 
likely not optimal but should not be too far from the optimal. 
Finally, the sixth one is obtained by combining (see below) the SF-template of length 380 and the one of length 3. 
Although we could not find a better SF-template with 7 colors, the last inequality is definetely very far from the 
optimal value. One may try to seek better templates by constraining less the search space and by using 
Monte-Carlo methods, as in \cite{Bouzy2015AnAP}. This could be the suject of a future work.

\par
The previous inequalities give new lower bounds for \(S(n)\) for
\( n \geqslant 9 \). We compute the lower
bounds for \( n \in [\![8,15]\!] \) using the four different inequalities, please notice that the best values for \( n
=8\) and \(n = 13\) were obtained thanks to the first one, found by Rowley. The best lower bounds are highlighted.

\begin{center}
\begin{tabular}{|*{5}{c|}}
    \hline
	\(n\) & 8 & 9 & 10 & 11 \\
	\hline
	\(33S(n-3) + 6 \) & \cellcolor{yellow} 5286 & 17694 & 55446 & 174444\\
	\hline
	\(111S(n-4) + 43 \) & 4927 & \cellcolor{yellow} 17803 & 59539 & 186523\\
	\hline
	\(380S(n-5) + 148 \) & 5088 & 16868 & \cellcolor{yellow} 60948 & \cellcolor{yellow} 203828 \\
	\hline
	\(1140S(n-6) + 528 \) & 5088 & 15348 & 50688 & 182928\\
	\hline
	\hline
	\(n\) & 12 & 13 & 14 & 15 \\
	\hline
	\(33S(n-3) + 6 \) & 587505 & \cellcolor{yellow} 2011290 & 6726330 & 21072090\\
	\hline
	\(111S(n-4) + 43 \) & 586789 & 1976176 & 6765271 & 22624951 \\
	\hline
\(380S(n-5) + 148 \) & \cellcolor{yellow} 638548 & 2008828 & \cellcolor{yellow} 6765288 & \cellcolor{yellow} 23160388 \\\hline
	\(1140S(n-6) + 528 \) & 611568 & 1915728 & 6026568 & 20295948 \\
	\hline
\end{tabular}
\end{center}

Except for 8, 9 and 13, the best lower bounds are obtained thanks to
the third inequality \( S(n+5) \geqslant 380S(n) + 148\). The table
doesn't go any further, but the same inequality allows to improve the
lower bounds for every \( n \geqslant 15 \).

\begin{corollary}
The growth rate for Schur numbers (and Ramsey numbers \(R_n(3)\))  satisfies \(\gamma \geqslant \sqrt[5]{380} \approx 3.28 \).
\end{corollary}

\begin{proof}[\textsc{Proof.}]
It is a mere consequence of the inequality \( S(n+5) \geqslant 380S(n) + 148\). As for Ramsey
numbers, the following inequality holds \(S(n) \leqslant R_n(3) - 2\) (see \cite{AbbottHanson}).
\end{proof}

### 2021-04-16 article/weakschur.tex
reorganisation

\begin{align*}
 \\\pi(\pi(x)+\pi(y)) & = (Id+a\mathds{1}_{ [\![0,b]\!]})(\pi(x)+\pi(y)\text{ mod a})\\
& = (Id+a\mathds{1}_{ [\![0,b]\!]})((Id+a\mathds{1}_{ [\![0,b]\!]})(x \mod a)+(Id+a\mathds{1}_{ [\![0,b]\!]})(y \mod a)\ mod a) \\
& = (Id+a\mathds{1}_{ [\![0,b]\!]})(((x \mod a) + (y \mod a)) \mod a)\\
& =(Id+a\mathds{1}_{ [\![0,b]\!]})( x+y \mod a)\\
& =\pi(x+y)
\end{align*}
\end{proof}

@@

\begin{remark}
In the following theorem, \(\lambda\) is the function which maps an element x to its line number, as mentioned in the intuitive explanation.
\end{remark}

@@

\begin{corollary}
\( \forall (n,k) \in (\mathbb{N}^*)^2 \text{, } let ~ b_{max}=max \{b\in \mathbb{N}^*/ \WS_b^+(n+1) = \WS^+(n+1)\},\) \[
\WS(n+k) \geqslant S(k) \WS^+(n+1)+b_{max}\]
\end{corollary}

\begin{remark}
In the SF-template construction for Schur numbers, the additive constant comes from the fact that the special color does
not necessarily appear right at the begining of the repeating pattern. Likewise, \(b_max\) can actually be replaced by \\
\[\max \left\{ b + \min (A_{n+1} \backslash [\![1, b ]\!]) - 1~|~ \WS_b^+(n+1) = \WS^+(n+1) \right\}\]
\end{remark}

\begin{remark}
Acutally, like for SF-templates, the additive constant of a WSF-template (in the form given by the above remark) can be improved
by weakening the hypotheses made on the last row. The principle behind it is the same as in the analogous proposition for SF-templates.
\end{remark}

@@

\begin{remark}
The WSF-templates can actually be fine-tuned further. However, it gives only minor improvements (most likely only additive
constant) at the cost of dramatically increasing the size of the search space while. Therefore, it does not seem relevant to
use this sophistications given that we could not even find good WSF-templates with 5 colors using a computer (here good
means better than those obtain by combining smaller templates).

@@

The first two inequalities were found by Rowley, they are detailed in [2]. The third inequality is optimal and was found with a SAT solver.
It uses the first sophistication explained in the previous subsection in order to add the last number in the first color.
As for the fourth inequality, it was obtained by combining an optimal SF-template of length 33 with a WSF-template of length 4.
The best template we could get with a computer search gives the inequality \(\WS(n+4) \geqslant 127 S(n) + 68\).
It was also found with the SAT solver. In order to reduce the search space, we only looked for WSF-templates of
5 colors which start with a good \(\WS(4)\) partition. However, this approach most likely prevents us from finding the best WSF-templates
as we explain in the next subsection for weakly sum-free partitions. We highly suspect that there exists more efficient WSF-templates
with \(n \geqslant 5\) colors. One may try to go over a different search space using a Monte-Carlo method, as in \cite{Bouzy2015AnAP}.
This could be the suject of a future work. Further details about the encoding as a SAT problem can be found in the
\hyperref[SAT]{SAT section}.

>>>>

\begin{align*}
 \\\pi(\pi(x)+\pi(y)) & = (Id+a\mathds{1}_{ [\![0,b]\!]})(\pi(x)+\pi(y) \mod a)\\
& = (Id+a\mathds{1}_{ [\![0,b]\!]})((Id+a\mathds{1}_{ [\![0,b]\!]})(x \mod a)+(Id+a\mathds{1}_{ [\![0,b]\!]})(y \mod a)\ mod a) \\
& = (Id+a\mathds{1}_{ [\![0,b]\!]})(((x \mod a) + (y \mod a)) \mod a)\\
& =(Id+a\mathds{1}_{ [\![0,b]\!]})( x+y \mod a)\\
& =\pi(x+y)
\end{align*}
\end{proof}

@@

\begin{remark}
In the following theorem, \(\lambda\) is the function which maps an element x to its line number as mentioned in the intuitive explanation.
\end{remark}

@@

\begin{corollary}
For every \((n,k) \in (\mathbb{N}^*)^2\), define \( b_{max} = \max \{b\in \mathbb{N}^*/ \WS_b^+(n+1) = \WS^+(n+1)\}\). 
Then the following inequality holds.
\[\WS(n+k) \geqslant S(k) \WS^+(n+1)+b_{max}\]
\end{corollary}

\begin{remark}
In the SF-template construction for Schur numbers, the additive constant comes from the fact that the special color does
not necessarily appear right at the begining of the repeating pattern. Likewise, \(b_{max}\) can actually be replaced by \\
\[\max \left\{ b + \min (A_{n+1} \backslash [\![1, b ]\!]) - 1~|~ \WS_b^+(n+1) = \WS^+(n+1) \right\}\]
\end{remark}

@@

\begin{remark}
Acutally, like for SF-templates, the additive constant of a WSF-template can be improved by weakening the hypotheses made on 
the last row. The principle behind it is the same as in the analogous proposition for SF-templates. 
\end{remark}

\begin{remark}
The WSF-templates can actually be fine-tuned further. However, it gives only minor improvements (most likely only an additive
constant) at the cost of dramatically increasing the size of the search space. Therefore, it does not seem relevant to
use this sophistications given that we could not even find good WSF-templates with 5 colors using a computer (here good
means better than those obtain by combining smaller templates).

@@

The first two inequalities were found by Rowley, they are detailed in [2]. The third inequality is optimal and was found with a SAT solver.
It uses the first sophistication explained in the previous subsection in order to add the last number in the first color.
As for the fourth inequality, it was obtained by combining an optimal SF-template of length 33 with a WSF-template of length 4.
The best template we could get with a computer search gives the inequality \(\WS(n+4) \geqslant 127 S(n) + 68\).
It was also found with the SAT solver. In order to reduce the search space, we only looked for WSF-templates of
5 colors which start with a good \(\WS(4)\) partition. However, this approach most likely prevents us from finding the best WSF-templates
as we explain in the next subsection for weakly sum-free partitions. We highly suspect that there exists more efficient WSF-templates
with \(n \geqslant 5\) colors. One may try to go over a different search space using a Monte-Carlo method, as in \cite{Bouzy2015AnAP}.
This could be the suject of a future work. Further details about the encoding as a SAT problem can be found in \cite{Heule2017}.
